---
title: "Seurat Trajectory Analyis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(Seurat)
library(dplyr)
library(tidyverse)
library(SeuratWrappers)
library(monocle3)
library(ggplot2)
```



# Reading the Seurat object

```{r,warning=FALSE,message=FALSE,echo=FALSE}
IPF_processed <- readRDS(r"(D:\A\App\processed_data.rds)")
```




```{r,warning=FALSE,message=FALSE}

identity_labels <- IPF_processed$orig.ident
unique_identity_labels <- unique(identity_labels)
print(unique_identity_labels)

cluster_labels <- IPF_processed$seurat_clusters
clus_identity_labels <- unique(cluster_labels)
print(clus_identity_labels)


IPF_processed@meta.data[["ident"]] <-IPF_processed@active.ident

```


# **Step 1: convert to seurat object to celll data object**


```{r,warning=FALSE,message=FALSE,echo=FALSE}
cds <- as.cell_data_set(IPF_processed,group.by = 'ident')
```



```{r,warning=FALSE,message=FALSE,echo=FALSE}
# changing the assay name 

DefaultAssay(IPF_processed) <- "RNA"


# setting  gene annotaion 
fData(cds)$gene_short_name <- rownames(fData(cds))

```



# **step 2 Cluster cells (using clustering info from seurat's UMAP)**


## Assing all the cells  to one  Partion
```{r,warning=FALSE,message=FALSE,echo=FALSE}
recreate.partion  <- c(rep(1,length(cds@colData@rownames)))
names(recreate.partion)  <- cds@colData@rownames
recreate.partion <- as.factor(recreate.partion)

cds@clusters$UMAP$partitions <- recreate.partion

```

## Assiging the cluster information
```{r,warning=FALSE,message=FALSE,echo=FALSE}
list_cluster <- IPF_processed@active.ident
cds@clusters$UMAP$clusters <- list_cluster
```


## Assign UMAP coordiantes - cell embeddings 

```{r,warning=FALSE,message=FALSE,echo=FALSE}
cds@int_colData@listData$reducedDims$UMAP  <- IPF_processed@reductions$umap@cell.embeddings
```



# **Plot  the cluster  before learning the trajectory**
```{r,warning=FALSE,message=FALSE}
cluster_no_before <- plot_cells(cds,color_cells_by = 'seurat_clusters',
                             label_groups_by_cluster = FALSE,
                             group_label_size = 5)+
  theme(legend.position = "right")

cluster_before <- plot_cells(cds,color_cells_by = 'ident',
           label_groups_by_cluster = FALSE,
           group_label_size = 5)+
  theme(legend.position = "right")

print(cluster_before)
print(cluster_no_before)
```

# **step3 Learning the Trajectory **

```{r,warning=FALSE,message=FALSE,echo=FALSE}
cds <- learn_graph(cds, use_partition = FALSE)
```



# **Plot  the cluster After learning the trajectory**
```{r,warning=FALSE,message=FALSE}
# plot after learning 
cluster_no_after<- plot_cells(cds,color_cells_by = 'seurat_clusters',
                                label_groups_by_cluster = FALSE,
                                group_label_size = 5)+
  theme(legend.position = "right")

cluster_after <- plot_cells(cds,color_cells_by = 'ident',
                             label_groups_by_cluster = FALSE,
                             group_label_size = 5)+
  theme(legend.position = "right")

print(cluster_no_after)
print(cluster_after)
```


```{r,warning=FALSE,message=FALSE,echo=FALSE}
# getting the unique cluster names 
cluster_names <- as.character(clusters(cds))

unique_clusters <- unique(cluster_names)
print(unique_clusters)
```

# **Step3 Order the cells based on the pseudotime  analysis**

```{r,warning=FALSE,message=FALSE}
#setting the root to get the pseudotime , here  ["Adventitia"] as taken root

cds <- order_cells(cds,reduction_method = 'UMAP',root_cells = colnames(cds[,clusters(cds) == "Alveolar"]))
```


# **Plot based on the oreder of cells of Pseudotime**


```{r,warning=FALSE,message=FALSE,echo=FALSE}

a<-plot_cells(cds,color_cells_by = 'pseudotime',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           group_label_size = 5)+
  theme(legend.position = "right")

b <-plot_cells(cds,color_cells_by = 'pseudotime',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           group_label_size = 5)+
  theme(legend.position = "right")
print(a)
print(b)
```


# **step4 Cells oredered by monocle3 pseudotime and visulaize the range of pseudotime**

```{r,warning=FALSE,message=FALSE,echo=FALSE}

cds$molocle3_pseudotime <- pseudotime(cds)
data.pseudo <- as.data.frame(colData(cds))
```


# **Box plot based on the Median value of Pseudotime**

```{r,warning=FALSE,message=FALSE,echo=FALSE}
boxplot <- ggplot(data.pseudo, aes(x = molocle3_pseudotime, y = reorder(ident, molocle3_pseudotime, FUN = median), fill = ident)) +
  geom_boxplot() +
  labs(title = "Boxplot of molocle3_pseudotime by ident",
       x = "molocle3_pseudotime",
       y = "ident") +
  theme_minimal()


print(boxplot)
```


# **step 4 To find the genes that expressed as the moves in the trajectory **

```{r,warning=FALSE,message=FALSE,echo=FALSE}
cds <- estimate_size_factors(cds)
## Add gene names into CDS
cds@rowRanges@elementMetadata@listData[["gene_short_name"]] <- rownames(cds[["RNA"]])

Differenial_expressed_genes<- graph_test(cds,neighbor_graph = "principal_graph",cores = 3)

#View(Differenial_expressed_genes)

differential_genes_filtered <-Differenial_expressed_genes %>%
  arrange(q_value) %>%
  filter(status=="OK") 
  
```
```{r}
head(differential_genes_filtered)
```

# **Ploting the Diffrential Expressed  genes 

```{r,warning=FALSE,message=FALSE,echo=FALSE}
FeaturePlot(IPF_processed,features = c("C1QC","ARHGAP30","CCM2L","SOX7"))
```

#**visuaalizion pseudotime in seurat**

```{r,warning=FALSE,message=FALSE,echo=FALSE}
IPF_processed$pseudotime <- pseudotime(cds)
FeaturePlot(IPF_processed,features ="pseudotime",,label = T )
```



